Using Stochastic Modeling to Predict Long Timescale Transport Behavior of Solutes in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-08-26 Ben_Manuscripts/stochastic_transport/Outline.tex
Added back dummy energy minimization

%MRS: better to have the first outline shorter to iterate more quickly!

@@

@@

  Unfortunately, the timescales that we can simulate with MD are insufficient to be
  able to make well-converged predictions of macroscopic transport properties 
  traditionally used to characterize membranes in the lab. \\

@@

@@

  \noindent Three common mathematical models for modeling anomalous subdiffusion
  processes include continuous time random walks (CTRW), fractional Brownian motion
  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
  \begin{itemize}
    \item FBM is common in crowded, viscoelastic environments where each step comes 
    from a Gaussian distribution but is anti-correlated to its previous 
    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
    \item A CTRW is characterized by a distribution of hop lengths and 
    dwell times, where each each step is characterized by independent random draws from 
    each distribution.\cite{montroll_random_1965,morrin_three_2018}
    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
    \item The processes described above can happen alone or in combination.  	
  \end{itemize}
  %MRS: talk about which of these should apply given the data mechanistically, as well as fit.
  %BJC1: I believe that's what I try to do in the paragraph below. I can be a little more explicit

  % BJC: Should this be in the results? Lots of figures to show in support of the claims below
  %MRS: can you provide mechanistic reasoning for the below? Then can show the claims for the reasonable possibilities.

  \noindent We believe that solutes in the system studied here exhibit subordinated 
  fractional Brownian motion (sFBM) where the parent process is FBM and the 
  leading process is a CTRW. 
  \begin{itemize}
  	\item The ensemble-averaged MSD differs from the time-averaged MSD, which
  	is indicative of non-ergodicity, a trait inherent to CTRWs but not FBM or RWFs.~\cite{thiel_weak_2014}  % BJC: need a figure to show this
  	\item We also observe non-stationary $z$-coordinate traces of each solute's
  	center of mass (COM). %BJC: figure of a z-coordinate trace in main text or in supporting info
  	\item For a pure CTRW, the time-averaged MSD should be linear.
  	~\cite{neusius_subdiffusion_2008,meroz_subdiffusion_2010}
  	\item However, a typical time-averaged solute MSD is sublinear (see supporting
  	information), which suggests that there is another underlying subdiffusive mechanism.
  	\item The hop lengths recorded after each dwell period are anti-correlated (See supporting information)
  	\item Given the viscoelastic nature of the monomers in our system, we believe
  	the hop lengths can be modeled with FBM. 
 	\item For subordinated FBM, it can be shown that
  	\begin{equation}
  	\langle x^2(t) \rangle \simeq t^{\alpha\beta}
  	\end{equation}
  	where $\alpha$ is the anomalous exponent characteristic of the leading CTRW process
  	and $\beta$ is the anomalous exponent characteristic of the parent FBM process. 
  \end{itemize}

>>>>

%MRS: better to have the first outline shorter to iterate more quickly!
%BJC1: right, this one has been in the works for a while now. Wrote the anomalous diffusion methods a long time ago

@@

  Unfortunately, the timescales that we can simulate with MD are insufficient to be
  able to make well-converged predictions of macroscopic transport properties 
  traditionally used to characterize membranes in the lab. \\
  \begin{itemize}
    \item However, if we use descriptive stochastic models that can capture solute
    dynamics, then we could project long timescale behavior in addition to gaining
    a deeper understanding of solute behavior on short timescales.
  \end{itemize}

  \noindent Researchers have developed a rigorous theoretical foundation for describing
  the motion of particles exhibiting non-Brownian, or anomalous, transport behavior.
  \begin{itemize}
    \item The tools introduced by fractional calculus are instrumental to this theory. 
    \item They allow us to generalize the normally linear diffusion equation using fractional
    derivative orders, providing descriptions of a much more diverse set of behavior. % diverging from the relatively simple Brownian model.
%    \item They facilitate proof of the relationship between the observed growth of a
%    particle's mean squared displacement and underlying dwell times 
%    \item In the absence of an external potential or complicated boundary conditions, 
%    one can define random walk based on independent hop length and dwell time 
%    distributions and arrive at any type of diffusive behavior. 
  \end{itemize}

  % BJC1: moved from methods
  \noindent Three well-known classes of behavior leading to anomalous subdiffusion are 
  continuous time random walks (CTRW), fractional Brownian motion
  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
  \begin{itemize}
    \item A CTRW is characterized by a distribution of hop lengths and 
    dwell times, where each each hop is characterized by independent random draws from 
    each distribution.\cite{montroll_random_1965,morrin_three_2018}
    \item FBM is common in crowded, viscoelastic environments where each jump comes 
    from a Gaussian distribution but is anti-correlated to its previous 
    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
    \item The processes described above can happen alone or in combination. 
    \item Meroz and Sokolov suggested a decision tree for determining which subdiffusive
    mechanisms apply to a system, adapted in Figure X. 
    \item Of course, more complicated variations of these models can arise which we 
    will incorporate as needed.	  % BJC1: e.g. fractional levy motion or langevin with levy noise
  \end{itemize}

  % BJC1: Flow chart here

  While elegant, pure anomalous diffusion theory may in some cases give only a qualitative
  understanding of transport mechanisms.
  \begin{itemize}
    \item Complex temporal relationships in between hops may have a significant effect on MSD
    \item Inhomogeneity of a medium or multiple types of interactions may lead to a number of
    distinct diffusive regimes. 
    \item In this case, a state based model may be useful. 
  \end{itemize}

  We can use the observations of transport from our previous work in order to create 
  state-based models.
  \begin{itemize}
  	\item We know that dynamics differ between solutes moving inside the pores and in the tails.
  	\item We can fit the above models to our data as a function of location.
  	\item Alternatively, we can abandon the anomalous diffusion theory and adapt the 
  	methods of Markov state models.
  \end{itemize}

  Markov state models blah blah blah.
  \begin{itemize}
    \item Markov state models are popular with researchers of proteins which have 
    many dynamical states occurring over multiple time scales.
  \end{itemize}

  We will extend the Markov state models framework to include dynamics between jumps.

@@

@@

  % BJC: quote from anomalous diffusion book:
  % "While the continuous time random walk model for Levy flights ... is a
  % conventient description, in all other cases the fraction Fokker-Planck or,
  % equivalently, the Langevin equation with white Levy stable noise are the 
  % description of choice" -- pretty sure this is exactly what I'm suggesting above.

@@

  % BJC1: moved this to intro
%  \noindent Three common mathematical models for modeling anomalous subdiffusion
%  processes include continuous time random walks (CTRW), fractional Brownian motion
%  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
%  \begin{itemize}
%    \item FBM is common in crowded, viscoelastic environments where each step comes 
%    from a Gaussian distribution but is anti-correlated to its previous 
%    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
%    \item A CTRW is characterized by a distribution of hop lengths and 
%    dwell times, where each each step is characterized by independent random draws from 
%    each distribution.\cite{montroll_random_1965,morrin_three_2018}
%    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
%    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
%    \item The processes described above can happen alone or in combination.  	
%  \end{itemize}
  %MRS: talk about which of these should apply given the data mechanistically, as well as fit.
  %BJC1: I believe that's what I try to do in the paragraph below. I can be a little more explicit

  % BJC: Should this be in the results? Lots of figures to show in support of the claims below
  %MRS: can you provide mechanistic reasoning for the below? Then can show the claims for the reasonable possibilities.
  \noindent We believe that solutes in the system studied here exhibit subordinated 
  fractional Brownian motion (sFBM) where the parent process is FBM and the 
  leading process is a CTRW. 
  \begin{itemize}
  	\item The ensemble-averaged MSD differs from the time-averaged MSD, which
  	is indicative of non-ergodicity, a trait inherent to CTRWs but not FBM or RWFs.~\cite{thiel_weak_2014}  % BJC: need a figure to show this
  	\item We also observe non-stationary $z$-coordinate traces of each solute's
  	center of mass (COM). %BJC: figure of a z-coordinate trace in main text or in supporting info
  	\item For a pure CTRW, the time-averaged MSD should be linear.
  	~\cite{neusius_subdiffusion_2008,meroz_subdiffusion_2010}
  	\item However, a typical time-averaged solute MSD is sublinear (see supporting
  	information), which suggests that there is another underlying subdiffusive mechanism.
  	\item The hop lengths recorded after each dwell period are anti-correlated (See supporting information)
  	\item Given the viscoelastic nature of the monomers in our system, we believe
  	the hop lengths can be modeled with FBM. 
 	\item For subordinated FBM, it can be shown that
  	\begin{equation}
  	\langle x^2(t) \rangle \simeq t^{\alpha\beta}
  	\end{equation}
  	where $\alpha$ is the anomalous exponent characteristic of the leading CTRW process
  	and $\beta$ is the anomalous exponent characteristic of the parent FBM process. 
  \end{itemize}
